On the Functionality of the Multilayer Perceptron

### 2020-11-04 Thesis/Drafts/Buell-PHYS799-ThesisDraft0.tex

\paragraph*{}Other than inspiration from the brain, Neural Networks are name \textit{networks} because of their nested composition nature. The model can be represented by a linear or acyclic computational graph which successively maps exactly how the repeated composition is structured. In a \textit{feed-forward network}, information is passes successively in one direction. For example functions could be chained together such as \cite{Goodfellow}
\begin{equation}
\label{eqn-FunctionChain}
F(x) = f^{(L-1)}\big[f^{(L-2)}\big[...f^{(1)}\big[f^{(0)}\big[x]\big]\big]...\big]\big]
\end{equation}
Where each function $f^{(l)}$ represents a layer of the network model. The number of layers in a neural network is referred to as the network \textit{depth}. The dimensionality of each layer is referred to as the network \textit{width} \cite{Geron,Loy}.

@@

\paragraph*{}The model is recursive by nature, the activations from one layer, $l-1$, are used to directly produce the activations of the next successive layer $l$. Thus eqn. (\ref{eqn-LayerFunction}) can be alternatively written as:
\begin{equation}
\label{eqn-altLayerFunction}
f^{(l)} : x^{(l-1)} \in \mathbb{R} \rightarrow x^{(l)} \in \mathbb{R}
\end{equation}
The array of activations, $x^{(0)}$, is the raw input given the neural network, called \textit{input} activations. Conversely, $x^{(L-1)}$ is referred to as\textit{output} activations \cite{Geron,James,Loy}.

\begin{algorithm}[H]
\caption{Forward propagation system in a standard deep neural network. Each iteration in the main \textit{for-loop} represents the execution of a layer, and passing the result to the "next" layer function. We assume each layter to follow a two-step structure, being (i) a linear transformation as in eqn. (\ref{eqn-LinearTransform}) and (ii) an element-wise non-linear activation function as in eqn. (\ref{eqn-elementActivation}).  A practical application of this algorithm should include batches of samples instead of a single sample.
\textcolor{red}{I Want to include this algorithm for context - especially on the programming end, but am not sure where to put it. }}
\label{algFeedForward}

@@

If training the network, the intermediate layer activations must be stored for back-propagation. Otherwise, they must be discarded.
$Z = \{\} \leftarrow$ array to hold pre-activations \\
$X = \{\} \leftarrow$ array to hold final activations \\

\FOR {$l = 1,2,...,L-2,L-1$}	
	\item Evaluate Linear Transformation
	\item $z^{(l)} \leftarrow$ 
	\item Apply non-linear Activation Function
	\item $x^{(l)} \leftarrow \sigma^{(l)} \big[ z^{(l)} \big]$ 
	\item Store values for later use in optimizer, otherwise can be discarded.
	\item $Z \leftarrow Z.add(z^{(l)})$
	\item $X \leftarrow X.add(x^{(l)})$
\ENDFOR

\item Network final prediction is last object in 'act' array.
\item $y^* \leftarrow x^{(L-1)} = X[-1]$
\item Return the pre-nonlinear activations and final-nonlinear activations.\\
\item The final index of $X$ represents the neural network's final prediction. \\
\RETURN $Z , X$

@@

\paragraph*{}As stated previously, a neural network is composed of a series of functions that are called successively to transform features (an input) into a prediction (an output). As shown in eqn. (\ref{eqn-altLayerFunction}), each function feeds directly into the next as to form a sort of recursive computational graph \cite{Goodfellow}.

\paragraph*{}Typically, a layer function can be divided into two portions: (i.) a Linear transformation, with a bias addition, and (ii.) an element-wise activation transformation. This activation function typically is a non-linear function which allows for the modeling of increasingly complex decision-boundaries. The Linear transformation can be given by eqn (\ref{eqn-LinearTransform}) in the case of a 1D Dense layer (\ref{subsubsec-DenseLayer}) or eqn. (\ref{eqn-LinearTransform2}) in the case of a 2D Convolution layer.
\begin{equation}
\label{eqn-LinearTransform}
z^{(l)} = W^{(l)} x^{(l-1)} + b^{(l)}
\end{equation}
\begin{equation}
\label{eqn-LinearTransform2}
z^{(l)} = \big(W^{(l)} * x^{(l-1)}\big) + b^{(l)}
\end{equation} 
Where $W^{(l)} x^{(l-1)}$ denotes a matrix-vector product, and $\big(W^{(l)} * x^{(l-1)}\big)$ denotes a convolution product. In both cases, $W^{(l)}$ is the \textit{weighting-matrix} or \textit{weighting-kernel} for layer $l$, $b^{(l)}$ is the \textit{bias-vector} for layer $l$, $z^{(l)}$ are the \textit{linear-activations} for layer $l$ and $x^{(l-1)}$ is the final activations for layer $l-1$.

@@

\paragraph*{}The combination of eqn. (\ref{eqn-LinearTransform}) and eqn. (\ref{eqn-elementActivation}), for a layer $l$ , make up the function represented by that layer, as in eqn. (\ref{eqn-altLayerFunction}). Below, we discuss and describe the types of layer functions that are used to produce the classification model in this project.

@@

\paragraph*{}The Linear Dense Layer, often just called a \textit{Dense} Layer, or \textit{Fully-Connected} layer for short, was one of the earliest function types used in artificial neural network models \cite{Geron2,Loy}. They are among the most commonly used layer types. A dense layer is composed of a layer of \textit{artifical neurons}, each of which holds a numerical value within it, called the \textit{activation} of that neuron. This idea was developed back from McCulloch and Pitts' work \cite{McCulloch}, and was expanded upon by Frank Rosenblatt in  1957 \cite{Geron}.

\paragraph*{}We model a layer of neurons as a vector of floating-point numbers. Typically, it is required that the array be one-dimensional. Suppose a layer $(l)$ contains $n$ artificial neurons. We denote the array that hold those activations as $x^{(l)}$ and is given by:
\begin{equation}
\label{layer-DenseNeurons}
\vec{x}^{(l)} = \Big[ x_0, x_1, x_2, \hdots , x_{n-2}, x_{n-1} \Big]^T
\end{equation}
The activation of each entry is given by a linear-combination of activations from the previous layer, as outlined in eqn.(\ref{eqn-LinearTransform}) and eqn.(\ref{eqn-elementActivation}).

\paragraph*{}Suppose the layer $l-1$ contains $m$ neurons. Then the weighting-matrix, $W^{(l)}$ has shape $m \times n$, the bias-vector $b^{(l)}$ has shape $m \times 1$. We can denote the function in a similar manner to eqn.(\ref{eqn-altLayerFunction}) as:
\begin{equation}
\label{eqn-FunctionDense}
f^{(l)} : x^{(l-1)} \in \mathbb{R}^{(m \times 1)} \rightarrow x^{(l)} \in \mathbb{R}^{(n \times 1)}
\end{equation}

Thus for a dense layer $l$, the exact values of each activation is given by \cite{Geron,Loy}:
\begin{equation}
\begin{bmatrix}
x_0 \\ x_1 \\ \vdots \\ x_{n-1}
\end{bmatrix}^{(l)} =
\sigma^{(l)} \Bigg(
\begin{bmatrix}
w_{0,0} & w_{0,1} & \hdots & w_{0,m-1} \\
w_{1,0} & w_{1,1} & \hdots & w_{1,m-1} \\
\vdots & \vdots & \ddots & \vdots  \\
w_{n-1,0} & w_{n-1,1} & \hdots & w_{n-1,m-1} 
\end{bmatrix}^{(l)} 
\begin{bmatrix}
x_0 \\ x_1 \\ \vdots \\ x_{m-1}
\end{bmatrix}^{(l-1)} +
\begin{bmatrix}
b_0 \\ b_1 \\ \vdots \\ b_{n-1}
\end{bmatrix}^{(l)} \Bigg)
\end{equation}
or more compactly:
\begin{equation}
\label{eqn-DenseFeedForward}
x^{(l)} = \sigma^{(l)} \Big( W^{(l)} x^{(l-1)} + b^{(l)} \Big)
\end{equation}
Eqn. (\ref{eqn-DenseFeedForward}) is generally referred to as the \textit{dense layer feed-forward equation} \cite{Goodfellow}.

@@

\paragraph*{}Convolution layers emerged from studying the brain's visual cortex, and have been used from image recognition related tasks for around 40 years. \cite{Geron,Loy}). This layer get's it's name from it's implementation of the mathematical \textit{convolution} function. The 2D discrete-convolution of function or 2D array $A[i,j]$ and $B[i,j]$ is given by \cite{Goodfellow}:
\begin{equation}
\label{eqn-convolution}
C[i,j] = (A * B)[i,j] = \sum_{u}\sum_{v} A[i,j]B[i - u,j - v]
\end{equation}
Note that convolution is commutative: $(A * B) = (B * A)$. We implement a convolutional layer $f^{(l)})$, by creating a series of $K$ weighting matrices, called \textit{filters}, each of size $m \times n$, where $m,n \in \mathbb{Z}$. Often we choose $m = n$ and call the shape the \textit{convolution kernel size} \cite{Loy, Goodfellow}.

\paragraph*{}In the case of a 2D input array, $x^{(l-1)} \in \mathbb{R}^{(N \times M)}$, the convolutions filters "step" through the input data and repeatedly compute the element-wise product of each $m \times n$ weighting kernel, and the local activation of the $x^{(l-1)}$ array. The step size in each of the 2-dimension is known as the \textit{stride}. Each of the $K$ filters are used to generate a $m \times n$ feature map from the input activation. For an appropriately trained network, some filters may be dedicated to detecting vertical lines, horizontal lines, sharp edges, areas of mostly one color, etc. \cite{Geron,Loy}.

\paragraph*{}In general, for an $N \times M$ input, with kernel size $m \times n$, with stride size $1 \times 1$, and $K$ feature maps.
\begin{equation}
f^{(l)} : x^{(l-1)} \in \mathbb{R}^{(N \times M)} \rightarrow
x^{(l)} \in \mathbb{R}^{([N-n+1] \times [M-m+1] \times K)}
\end{equation}
As an example, consider a single filter map over input $x \in \mathbb{R}^{(4 \times 3)}$ and filter map $W \in \mathbb{R}^{(2 \times 2)}$ as shown in fig. (\ref{fig-2DConvExample}):
\begin{figure}[H]
\label{fig-2DConvExample}
\begin{center}
\begin{subfigure}{0.45\textwidth}
\includegraphics[scale=0.6]{../figures/2DConvExample(a)}
\caption{Input Activations}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
\includegraphics[scale=0.6]{../figures/2DConvExample(b)}
\caption{Convolution Filter}
\end{subfigure}
\begin{subfigure}{0.9\textwidth}
\centering
\includegraphics[scale=0.6]{../figures/2DConvExample(c)}
\caption{Convolution Result}
\end{subfigure}
\end{center}
\caption{The result of convolving an input (a) with an filter map (b) is a new set of activations (c). This Image was adapted from Goodfellow, pg. 325 \cite{Goodfellow}}
\end{figure}
Note that formal implementations include a bias vector and an activation function.

@@

\subsubsection{2-Dimensional Maximum Pooling Layer}

\paragraph*{}A Maximum Pooling layers simply returns the maximum neuron activation in a group of neurons. In the case of 2D Max Pooling, we choose a kernel size to be $m \times n$, just like in the 2D Convolution layer, and extract the maximum value in each window, and then step along accrdoing to the chosen stride size \cite{Loy,Goodfellow}.As an example, consider again figure (\ref{fig-2DConvExample}a). Using the $2 \times 2$ kernel on the $3 \times 4$ input, each box would then contain the maximum value of the input activations. We detail this in fig. (\ref{fig-2DMaxPool}):
\begin{figure}[h]
\label{fig-2DMaxPool}
\begin{center}
\includegraphics[scale=0.6]{../figures/2DConvExample(d)}
\end{center}
\caption{The result of 2D maximum-pooling an input array. This image was adapted from Loy, pg. 126 \cite{Loy}}
\end{figure}

@@

\subsubsection{1-Dimensional Flattening Layer}

\paragraph*{}A flattening layer is used to compress an array with two or more dimensions down into a single dimension. For a flattening layer $l$, multidimensional activations in layer $l-1$ are compressed down into a single dimensional array. We can use function notation to express this as:
\begin{equation}
\label{eqn-FlattenFunction}
f^{(l)} : x^{(l-1)} \in \mathbb{R}^{(M \times N \times ....)} \rightarrow
x^{(l)} \in \mathbb{R}^{(MN...\times 1)}
\end{equation}
The numerical value of each activation is left unchanged. For a layer with activation shape$N \times M$, the resulting activations are reshaped into $NM \times 1$ as shown in eqn (\ref{eqn-FlattenFunction}).

@@

\subsubsection{1-Dimensional Concatenation Layer}

\paragraph*{}The 1-Dimensional Concatenation Layer, also called 1D-Concat layer takes separate vectors of activations and combines them into a single 1D vector. Consider the layer activations $\vec{a}^{(l-1)}$ and $\vec{b}^{(l-1)}$ with shapes  $1 \times \alpha$ and $1 \times \beta$ respectively. They are the outputs of two different layers:
\begin{equation}
\vec{a}^{(l-1)} = \big[ a_{0},  a_{1}, ... ,  a_{\alpha-1} \big]
\text{\hspace{1cm} and \hspace{1cm}}
\vec{b}^{(l-1)} = \big[ b_{0},  b_{1}, ... ,  b_{\beta-1} \big]
\end{equation}
The result of the concatenation is a new 1D-array, $\vec{c}^{(l)}$ with size $1 \times \alpha\beta$:
\begin{equation}
\vec{c}^{(l)} = \big[ a_{0},  a_{1}, ... ,  a_{\alpha-1}, b_{0},  b_{1}, ... ,  b_{\beta-1} \big]
\end{equation}

\paragraph*{}We can denote this for $n$ arrays with function notation:
\begin{equation}
\label{eqn-ConcatenationFunction}
f^{(l)} : x^{(l-1)}_{a} \in \mathbb{R}^{(1 \times \alpha)} , x^{(l-1)}_{b} \in \mathbb{R}^{(1 \times \beta)} , ....  \rightarrow
x^{(l)}_{z} \in \mathbb{R}^{(1 \times \alpha\beta...)}
\end{equation}
The concatenation layer is used to combine activation non-adjacent, or non-related layers into a single new layer. In this case of the model used in the project, we use it to combine the activations that result from two distinct models into a single layer that is then transformed into a final output. This process is detailed further in sec. (\ref{sec-Architecture}).

@@

\paragraph*{}Activation functions are a key parameter in the behavior of neural networks. As discussed in sec. (\ref{sec-Training}) a layer function, $f^{(l)}$ is generally composed of a linear transformation, as in eqn. (\ref{eqn-LinearTransform}), and then an element-wise activation function as in eqn. (\ref{eqn-elementActivation}). It is this second step that allows for neural networks to model the incredibly complex decision boundaries that are found in real-world problems \cite{Geron,Loy}. In this section, we detail the activation functions used in out classification neural network.

@@

\paragraph*{}The Rectified Linear Unit (ReLU) activation function acts element-wise on the activations in a given layer. If the activation of a neurons is non-zero or non-negative, the value is untouched, otherwise a $0$ is returned. For an input activation $x$, ReLU is defined by:
\begin{equation}
\label{eqn-ReLU}
\text{ReLU}\big[x\big] = \max\big(0,x\big) = \left\{
\begin{array}{ll}
    x & \text{if\hspace{0.5cm}} x > 0 \\
    0 & \text{otherwise}
\end{array} 
\right.
\end{equation}
ReLU is only defined on a real domain. We provide a visualization of the function in fig. (\ref{fig-ReLU}).
\begin{figure}[H]
\label{fig-ReLU}
\begin{center}
\includegraphics[scale=0.2]{../Figures/RectifiedLinearUnit}
\end{center}
\caption{Rectifed Linear Unit (ReLU) activation function}
\end{figure}
For our chosen architecture, ReLU is applied to the activations in every single Convolution layer and Dense Layer, with the exception of the output dense layer.

@@

\subsubsection{Softmax}

\paragraph*{}The softmax activation function is commonly used in the output layer of a multi-class classification network, such as the one we implement. When softmax acts on a vector or activation, the result a non-negative output vector. with an $L_1$ norm of $1$ \cite{Geron,Goodfellow,Virtanen}. The $i$-th element in a softmax activation function is given by:
\begin{equation}
\label{eqn-Softmax}
\text{Softmax}\big[ x \big]_{i} = \frac{1}{\sum_{j}e^{x_j}}e^{x_i}
\end{equation}

\paragraph*{}This function is almost only used in the output layer of a neural network, and encodes a discrete categorical probability. For example, if eqn.(\ref{eqn-Softmax}) returns a vector of the form: $[ 0.75 , 0.25]$, we interpret this as a sample having a $75\%$ chance of beloinging to class $0$ and a $25\%$ change of belonging to class $1$.

@@

\paragraph*{}We consider the set of parameters $\Theta$ as a concatenation of each applicable layer's weighting matrix and bias vector such that:
\begin{equation}
\label{eqn-Theta}
\Theta = \big\{
W^{(0)}, b^{(0)}, W^{(1)}, b^{(1)}, W^{(2)}, b^{(2)}, ... ,W^{(L-2)}, b^{(L-2)}, W^{(L)}, b^{(L-1)},
\big\}
\end{equation}
Each parameter is floating-point number, which is stored in a weight matrix $W^{(l)}$ or bias vector $b^{(l)}$. Each value for each parameter contributes to the output of the function. Note that $W^{(l)}$ or $b^{(l)}$, itself is a parameter to train, but rather the entries that make up those structures must be trained. In increasingly complicated neural networks, there can be upwards of hundreds of thousands,or even millions of elements in $\theta$ making a neural network a functions that exist is a very high dimensional parameter-space \cite{Geron,Goodfellow}.

@@

>>>>

\paragraph*{}Other than inspiration from the brain, Neural Networks are name \textit{networks} because of their nested composition nature. The model can be represented by a linear computational graph which successively maps exactly how the repeated composition is structured. In a \textit{feed-forward network}, information is passed successively in one direction. For example functions could be chained together such as \cite{Goodfellow}
\begin{equation}
\label{eqn-FunctionChain}
F(x) = f^{(L-1)}\big[f^{(L-2)}\big[...f^{(1)}\big[f^{(0)}\big[x]\big]\big]...\big]\big]
\end{equation}
Where each function $f^{(l)}$ represents a layer of the network model. The number of layers in a neural network is referred to as the network \textit{depth}. The dimensionality of each layer is referred to as the network \textit{width} \cite{Geron,Loy}.

@@

\paragraph*{}The model is recursive by nature, the activations from one layer, $l-1$, are used to directly produce the activations of the next successive layer $l$. Thus Eq. (\ref{eqn-LayerFunction}) can be alternatively written as:
\begin{equation}
\label{eqn-altLayerFunction}
f^{(l)} : x^{(l-1)} \in \mathbb{R} \rightarrow x^{(l)} \in \mathbb{R}
\end{equation}
The array of activations, $x^{(0)}$, is the raw input given the neural network, called \textit{input} activations. Conversely, $x^{(L-1)}$ is referred to as \textit{output} activations \cite{Geron,James,Loy}.

\begin{algorithm}[H]
\caption{Forward propagation system in a standard deep neural network. Each iteration in the main \textit{for-loop} represents the execution of a layer, and passing the result to the "next" layer function. We assume each layer to follow a two-step structure, being (i) a linear transformation as in Eq. (\ref{eqn-LinearTransform}) and (ii) an element-wise non-linear activation function as in Eq. (\ref{eqn-elementActivation}).  A practical application of this algorithm should include batches of samples instead of a single sample.
\textcolor{red}{I Want to include this algorithm for context - especially on the programming end, but am not sure where to put it. }}
\label{algFeedForward}

@@

If training the network, the intermediate layer activations must be stored for back-propagation. Otherwise, they may be discarded. \\
$Z = \{\} \leftarrow$ array to hold pre-activations \\
$X = \{\} \leftarrow$ array to hold final activations \\

\FOR {$l = 1,2,...,L-2,L-1$}	
	\item Evaluate Linear Transformation
	\item $z^{(l)} \leftarrow W^{(l)} x^{(l-1)} + b^{(l)}$ 
	\item Apply non-linear Activation Function
	\item $x^{(l)} \leftarrow \sigma^{(l)} \big[ z^{(l)} \big]$ 
	\item Store values for later use in optimizer, otherwise can be discarded.
	\item $Z \leftarrow Z.add(z^{(l)})$
	\item $X \leftarrow X.add(x^{(l)})$
\ENDFOR

\item Network final prediction is last object in $X$ array.
\item $y^* \leftarrow x^{(L-1)} = X[-1]$
\item Return the pre-nonlinear activations and final-nonlinear activations.\\
\RETURN $Z , X$

@@

\paragraph*{}As stated previously, a neural network is composed of a series of functions that are called in succession to transform features (an input) into a prediction (an output). As shown in Eq. (\ref{eqn-altLayerFunction}), each function feeds directly into the next as to form a sort of recursive computational graph \cite{Goodfellow}.

\paragraph*{}Typically, a layer function can be divided into two portions: (i.) a Linear transformation, with a bias addition, and (ii.) an element-wise activation transformation. This activation function typically is a non-linear function which allows for the modeling of increasingly complex decision-boundaries. The Linear transformation can be given by Eq.(\ref{eqn-LinearTransform}) in the case of a 1D Dense layer (\ref{subsubsec-DenseLayer}) or Eq. (\ref{eqn-LinearTransform2}) in the case of a 2D Convolution layer.
\begin{equation}
\label{eqn-LinearTransform}
z^{(l)} = W^{(l)} x^{(l-1)} + b^{(l)}
\end{equation}
\begin{equation}
\label{eqn-LinearTransform2}
z^{(l)} = \big(W^{(l)} * x^{(l-1)}\big) + b^{(l)}
\end{equation} 
Where $W^{(l)} x^{(l-1)}$ denotes a matrix-vector product, and $\big(W^{(l)} * x^{(l-1)}\big)$ denotes a convolution product. In both cases, $W^{(l)}$ is the \textit{weighting-matrix} or \textit{weighting-kernel} for layer $l$, $b^{(l)}$ is the \textit{bias-vector} for layer $l$, $z^{(l)}$ are the \textit{linear-activations} for layer $l$ and $x^{(l-1)}$ is the final activations for layer $l-1$.

@@

\paragraph*{}The combination of Eq. (\ref{eqn-LinearTransform}) and Eq. (\ref{eqn-elementActivation}), for a layer $l$ , make up the function represented by that layer, as in Eq. (\ref{eqn-altLayerFunction}). Below, we discuss and describe the types of layer functions that are used to produce the classification model in this project.

@@

\paragraph*{}The Linear Dense Layer, often just called a \textit{Dense} Layer, or \textit{Fully-Connected} layer, was one of the earliest function types used in artificial neural network models \cite{Geron2,Loy,McCulloch}. They are among the most commonly used layer types. A dense layer is composed of a layer of \textit{artifical neurons}, each of which holds a numerical value within it, called the \textit{activation} of that neuron. This idea was developed from McCulloch and Pitts' work \cite{McCulloch} in attempting to model cognitive processes as mathematical functions, and was expanded upon by Frank Rosenblatt in  1957 \cite{Geron}.

\paragraph*{}We model a layer of neurons as a vector of floating-point numbers. Typically, it is required that the array be one-dimensional. Suppose a layer $(l)$ contains $n$ artificial neurons. We denote the array that hold those activations as $x^{(l)}$ and is given by:
\begin{equation}
\label{layer-DenseNeurons}
\vec{x}^{(l)} = \Big[ x_0, x_1, x_2, \hdots , x_{n-2}, x_{n-1} \Big]^T
\end{equation}
The activation of each entry is given by a linear-combination of activations from the previous layer, Eq.(\ref{eqn-LinearTransform}) and the transformation in Eq.(\ref{eqn-elementActivation}).

\paragraph*{}Suppose the layer $l-1$ contains $m$ neurons. Then the weighting-matrix, $W^{(l)}$ has shape $m \times n$, the bias-vector $b^{(l)}$ has shape $m \times 1$. We can denote the function in a similar manner to Eq.(\ref{eqn-altLayerFunction}) as:
\begin{equation}
\label{eqn-FunctionDense}
f^{(l)} : x^{(l-1)} \in \mathbb{R}^{(m \times 1)} \rightarrow x^{(l)} \in \mathbb{R}^{(n \times 1)}
\end{equation}

Thus for a dense layer $l$, the exact values of each activation is given by \cite{Geron,Loy}:
\begin{equation}
\begin{bmatrix}
x_0 \\ x_1 \\ \vdots \\ x_{n-1}
\end{bmatrix}^{(l)} =
\sigma^{(l)} \Bigg(
\begin{bmatrix}
w_{0,0} & w_{0,1} & \hdots & w_{0,m-1} \\
w_{1,0} & w_{1,1} & \hdots & w_{1,m-1} \\
\vdots & \vdots & \ddots & \vdots  \\
w_{n-1,0} & w_{n-1,1} & \hdots & w_{n-1,m-1} 
\end{bmatrix}^{(l)} 
\begin{bmatrix}
x_0 \\ x_1 \\ \vdots \\ x_{m-1}
\end{bmatrix}^{(l-1)} +
\begin{bmatrix}
b_0 \\ b_1 \\ \vdots \\ b_{n-1}
\end{bmatrix}^{(l)} \Bigg)
\end{equation}
or more compactly:
\begin{equation}
\label{eqn-DenseFeedForward}
x^{(l)} = \sigma^{(l)} \Big( W^{(l)} x^{(l-1)} + b^{(l)} \Big)
\end{equation}
Eq. (\ref{eqn-DenseFeedForward}) is generally referred to as the \textit{dense layer feed-forward equation} \cite{Goodfellow}. \textcolor{red}{How Do I finish this section? Uses of Dens layer, physical meaning?}

@@

\paragraph*{}Convolution layers emerged from studying the brain's visual cortex, and have been used from image recognition related tasks for around 40 years. \cite{Geron,Loy}). This layer get's it's name from it's implementation of the mathematical \textit{convolution} function. The 2D discrete-convolution of function or 2D array $A[i,j]$ and $B[i,j]$ is given by \cite{Goodfellow}:
\begin{equation}
\label{eqn-convolution}
C[i,j] = (A * B)[i,j] = \sum_{u}\sum_{v} A[i,j]B[i - u,j - v]
\end{equation}
Note that convolution is commutative: $(A * B) = (B * A)$. We implement a convolutional layer $f^{(l)}$ by creating a series of $K$ weighting matrices, called \textit{filters}, each of size $m \times n$, where $m,n \in \mathbb{Z}$. Often we choose $m = n$ and call the shape the \textit{convolution kernel size} \cite{Loy, Goodfellow}.

\paragraph*{}In the case of a 2D input array, $x^{(l-1)} \in \mathbb{R}^{(N \times M)}$, the convolutions filters "step" through the input data and repeatedly compute the element-wise product of each $m \times n$ weighting kernel, and the local activations of the $x^{(l-1)}$ array. The step size in each of the 2-dimension is known as the \textit{stride}. Each of the $K$ filters are used to generate a $m \times n$ feature map from the input activation. For an appropriately trained network, some filters may result in detecting vertical lines, horizontal lines, sharp edges, areas of mostly one color, etc. \cite{Geron,Loy}.

\paragraph*{}In general, for an $N \times M$ input, with kernel size $m \times n$, with stride size $1 \times 1$, and $K$ feature maps.
\begin{equation}
f^{(l)} : x^{(l-1)} \in \mathbb{R}^{(N \times M)} \rightarrow
x^{(l)} \in \mathbb{R}^{([N-n+1] \times [M-m+1] \times K)}
\end{equation}
As an example, consider a single filter map over input $x \in \mathbb{R}^{(4 \times 3)}$ and filter map $W \in \mathbb{R}^{(2 \times 2)}$ as shown in Fig. (\ref{fig-2DConvExample}):
\begin{figure}[H]
\begin{center}
\begin{subfigure}{0.45\textwidth}
\includegraphics[scale=0.6]{../figures/2DConvExample(a)}
\caption{Input Activations}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
\includegraphics[scale=0.6]{../figures/2DConvExample(b)}
\caption{Convolution Filter}
\end{subfigure}
\begin{subfigure}{0.9\textwidth}
\centering
\includegraphics[scale=0.6]{../figures/2DConvExample(c)}
\caption{Convolution Result}
\end{subfigure}
\end{center}
\caption{The result of convolving an input (a) with an filter map (b) is a new set of activations (c). This Image was adapted from Goodfellow, pg. 325 \cite{Goodfellow}}
\label{fig-2DConvExample}
\end{figure}
Note that formal implementations include a bias vector and an activation function.

@@

\subsubsection{2-Dimensional Maximum Pooling Layer}
\label{subsubsec-2DPool}

\paragraph*{}A Maximum Pooling layers simply returns the maximum neuron activation in a group of neurons. In the case of 2D Max Pooling, we choose a kernel size to be $m \times n$, just like in the 2D Convolution layer, and extract the maximum value in each window, and then step along accrdoing to the chosen stride size \cite{Loy,Goodfellow}.As an example, consider again figure (\ref{fig-2DConvExample}a). Using the $2 \times 2$ kernel on the $3 \times 4$ input, each box would then contain the maximum value of the input activations. We detail this in Fig. (\ref{fig-2DMaxPool}):
\begin{figure}[h]
\label{fig-2DMaxPool}
\begin{center}
\includegraphics[scale=0.6]{../figures/2DConvExample(d)}
\end{center}
\caption{The result of 2D maximum-pooling an input array. This image was adapted from Loy, pg. 126 \cite{Loy}}
\end{figure}

@@

\subsubsection{1-Dimensional Flattening Layer}
\label{subsubsec-1DFlatten}

\paragraph*{}A flattening layer is used to compress an array with two or more dimensions down into a single dimension. For a flattening layer $l$, multidimensional activations in layer $l-1$ are rearranged down into a single dimensional array. Note that this is not like projecting the data into a lower dimension, but rather is simple the reorganization of values into an array of a single axis. We can use function notation to express this as:
\begin{equation}
\label{eqn-FlattenFunction}
f^{(l)} : x^{(l-1)} \in \mathbb{R}^{(M \times N \times ....)} \rightarrow
x^{(l)} \in \mathbb{R}^{(MN...\times 1)}
\end{equation}
The numerical value of each activation is left unchanged. For a layer with activation shape$N \times M$, the resulting activations are reshaped into $NM \times 1$ as shown in Eq.(\ref{eqn-FlattenFunction}).

@@

\subsubsection{1-Dimensional Concatenation Layer}
\label{subsubsec-1DConcat}

\paragraph*{}The 1-Dimensional Concatenation Layer, also called 1D-Concat layer takes separate vectors of activations and combines them into a single 1D vector. Consider the layer activations $\vec{a}^{(l-1)}$ and $\vec{b}^{(l-1)}$ with shapes  $1 \times \alpha$ and $1 \times \beta$ respectively. They are the outputs of two different layers:
\begin{equation}
\vec{a}^{(l-1)} = \big[ a_{0},  a_{1}, ... ,  a_{\alpha-1} \big]
\text{\hspace{1cm} and \hspace{1cm}}
\vec{b}^{(l-1)} = \big[ b_{0},  b_{1}, ... ,  b_{\beta-1} \big]
\end{equation}
The result of the concatenation is a new 1D-array, $\vec{c}^{(l)}$ with size $1 \times \alpha + \beta$:
\begin{equation}
\vec{c}^{(l)} = \big[ a_{0},  a_{1}, ... ,  a_{\alpha-1}, b_{0},  b_{1}, ... ,  b_{\beta-1} \big]
\end{equation}

\paragraph*{}We can denote this for $n$ arrays with function notation:
\begin{equation}
\label{eqn-ConcatenationFunction}
f^{(l)} : x^{(l-1)}_{a} \in \mathbb{R}^{(1 \times \alpha)} , x^{(l-1)}_{b} \in \mathbb{R}^{(1 \times \beta)} , ....  \rightarrow
x^{(l)}_{z} \in \mathbb{R}^{(1 \times \alpha + \beta + ...)}
\end{equation}
The concatenation layer is used to combine activation non-adjacent, or non-related layers into a single new layer. In this case of the model used in the project, we use it to combine the activations that result from two distinct models into a single layer that is then transformed into a final output. This process is detailed further in sec. (\ref{sec-Architecture}).

@@

\paragraph*{}Activation functions are a key parameter in the behavior of neural networks. As discussed in sec. (\ref{sec-Training}) a layer function, $f^{(l)}$ is generally composed of a linear transformation, as in Eq. (\ref{eqn-LinearTransform}), and then an element-wise activation function as in Eq. (\ref{eqn-elementActivation}). It is this second step that allows for neural networks to model the incredibly complex decision boundaries that are found in real-world problems \cite{Geron,Loy}. In this section, we detail the activation functions used in out classification neural network.

@@

\paragraph*{}The Rectified Linear Unit (ReLU) activation function acts element-wise on the activations in a given layer. If the activation of a neurons is non-zero or non-negative, the value is untouched, otherwise a $0$ is returned. For an input activation $x$, ReLU is defined by:
\begin{equation}
\label{eqn-ReLU}
\text{ReLU}\big[x\big] = \max\big(0,x\big) = \left\{
\begin{array}{ll}
    x & \text{if\hspace{0.5cm}} x > 0 \\
    0 & \text{otherwise}
\end{array} 
\right.
\end{equation}
ReLU is only defined on a real domain. We provide a visualization of the function in Fig. (\ref{fig-ReLU}).
\begin{figure}[H]
\begin{center}
\includegraphics[scale=0.2]{../Figures/RectifiedLinearUnit}
\end{center}
\caption{Rectifed Linear Unit (ReLU) activation function}
\label{fig-ReLU}
\end{figure}
For our chosen architecture, ReLU is applied to the activations in every single Convolution layer and Dense Layer, with the exception of the output dense layer.

@@

\subsubsection{Softmax}
\label{subsubsec-Softmax}

\paragraph*{}The softmax activation function is commonly used in the output layer of a multi-class classification network, such as the one we implement. When softmax acts on a vector or activation, the result a non-negative output vector. with an $L_1$ norm of $1$ \cite{Geron,Goodfellow,Virtanen}. The $i$-th element in a softmax activation function is given by:
\begin{equation}
\label{eqn-Softmax}
\text{Softmax}\big[ x \big]_{i} = \frac{e^{x_i}}{\sum_{j}e^{x_j}}
\end{equation}

\paragraph*{}This function is almost only used in the output layer of a neural network, and encodes a discrete categorical probability. For example, if Eq.(\ref{eqn-Softmax}) returns a vector of the form: $[ 0.75 , 0.25]$, we interpret this as a sample having a $75\%$ chance of beloinging to class $0$ and a $25\%$ change of belonging to class $1$.

@@

\paragraph*{}We consider the set of parameters $\Theta$ as a concatenation of each applicable layer's weighting matrix and bias vector such that:
\begin{equation}
\label{eqn-Theta}
\Theta = \big\{
W^{(0)}, b^{(0)}, W^{(1)}, b^{(1)}, W^{(2)}, b^{(2)}, ... ,W^{(L-2)}, b^{(L-2)}, W^{(L)}, b^{(L-1)},
\big\}
\end{equation}
Each parameter is floating-point number, which is stored in a weight matrix $W^{(l)}$ or bias vector $b^{(l)}$. Each value for each parameter contributes to the output of the function. Note that $W^{(l)}$ or $b^{(l)}$, itself is a parameter to train, but rather the entries that make up those structures must be trained. Certain layers, such as pooling layers (\ref{subsubsec-2DPooling}), concatenation layers (\ref{subsubsec-1DConcat}), or flattening layers (\ref{subsubsec-1DFlatten}) do not contain any trainable parameters. Similarly, activation functions themselves are static and are not included in the list of parameters.

\paragraph*{}In increasingly complicated neural networks, there can be upwards of hundreds of thousands,or even millions of elements in $\theta$ making a neural network a functions that exist is a very high dimensional parameter-space \cite{Geron,Goodfellow,Levine}.

@@

\paragraph*{}Standard gradient-based learning, as given by the parameter update rule in Eq. (\ref{eqn-GradientLearning}) is cumbersome and often prone to numerical error. This includes phenomena where very small changes can compound and produce undesired effects such as \textit{vanishing} or \textit{exploding} gradients \cite{Geron,Goodfellow,Loy}. To combat this, many gradient-based optimizers have been developed over the years to mitigate some f these vulnerabilities.
